TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages

### 2021-01-22 article.tex
Paper TTT: doy otra pasada a la introducci'on

\en{Knowing how individual skills change over time is essential in sports and educational contexts.}
\es{Conocer c\'omo cambian las habilidades individuales a lo largo del tiempo es esencial para en contextos deportivos y educativos.}
%
\en{Since they are hidden variables, the best we can do is to estimate them based on their direct observable consequences: the outcome of problem solving and competitions.}
\es{Dado que son variables ocultas, lo mejor que podemos hacer es estimarlas a partir de sus consecuencias observables directas: el producto de resoluci\'on de problemas y competencias.}
%
\en{Considering only the frequency of positive results as an indicator of the individuals' ability could lead to wrong approximations, mainly because the outcome also depends on the difficulty of the challenge.}
\es{Considerar s\'olo la frecuencia de resultados positivos como indicador de la habilidad de los individuos puede conducir a aproximaciones erroneas, fundamentalmente porque su valor depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{For this reason, all widely used skill estimators are based on pairwise comparisons.}
\es{Por esta raz\'on, todos los estimadores de habilidad ampliamente usados se basan en comparaciones por pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{Thurstone1927} and~\cite{Zermelo1929}, it is assumed that the probability of the observed result $r$ depends on the performance $p$ of the agent $i$ and their opponent $j$, expressed as $P(\, r \,|\, p_i, \, p_j \,)$.}
\es{Desde los primeros modelos generativos, propuestos hace casi un siglo por~\cite{Thurstone1927} y~\cite{Zermelo1929}, se supone que la probabilidad de un resultado observado, $r$, depende del rendimiento, $p$, del agente $i$ y de su oponente $j$, expresada como $P(\, r \,|\, p_i, \, p_j \,)$.}
%
\en{The field continued to progress with the work of~\cite{Bradley1952} and~\cite{Mosteller1951a,Mosteller1951b,Mosteller1951c}, leading to the breakthrough that took place with the methodology developed by~\cite{Elo2008} for the US Chess Federation (USCF), which is still used by the International Chess Federation (FIDE).}
\es{El campo sigui\'o progresando con los trabajos de \cite{Bradley1952} y~\cite{Mosteller1951a,Mosteller1951b,Mosteller1951c}, que condujo al gran avance que tuvo lugar con la metodología desarrollada por~\cite{Elo2008} para la Federaci\'on de Ajedrez de los Estados Unidos (USCF), adoptada hasta el d\'ia de hoy por la Federaci\'on Internacional de Ajedrez (FIDE).}

@@

\en{All currently widely used skill estimators share some variant of the probabilistic model proposed by Elo~\citep{Glickman1999,Herbrich2007,VanDerLinden2016,Fox2010}.}
\es{Todos los estimadores de habilidad ampliamente utilizados actualmente comparten alguna variante del modelo probabil\'istico propuesto por Elo~\cite{Glickman1999, Herbrich2007, VanDerLinden2016, Fox2010}.}
%
\begin{figure}[h!]
\centering \small
    \tikz{         
    \node[det, fill=black!10] (r) {$r$} ; 
    \node[const, left=of r, xshift=-1.35cm] (r_name) {\small \en{Result}\es{Resultado}:}; 
    \node[const, right=of r] (dr) {\normalsize $ r = (d > 0)$};

@@

    \edge {d} {r};
    \edge {p1,p2} {d};
    \edge {s1} {p1};
    \edge {s2} {p2};
    %\node[invisible, right=of p2, xshift=4.35cm] (s-dist) {};
}
     \caption{
     \en{Generative model in which skills cause the observable results mediated by the difference of hidden performances, $d =p_i - p_j$, both random variables around their unknown true skill, $p \sim \N(s,\beta^2)$.}
    \es{Modelo generativo en el que las habilidades causan los resultados observables a trav\'es de la diferencia de rendimientos ocultos, $d=p_i-p_j$, ambas variables aleatorias centradas en la verdadera habilidad, $p \sim \N(s,\beta^2)$}
    %
    \en{The one with the highest performance wins, $r = (d > 0)$.}
    \es{Gana quien haya obtenido mayor rendimiento, $r = (d > 0)$.}
    %
    \en{Observable variables are painted gray, hidden in white, and constants are shown as black dots.}
    \es{Las variables observables se pintan de gris, la ocultas en blanco, y las constantes se muestran como puntos negros.}
    }
    \label{fig:generative_model}
\end{figure}
%
\en{Figure~\ref{fig:generative_model} provides a causal model in which skills generates the observable result.}
\es{La figura~\ref{fig:generative_model} ofrece un modelo causal en la que las habilidades generan el resultado observable.}
%
\en{The agents exhibit different performances at each event, varying around their true skill, $\N(p_i\,|\,s_i,\beta^2)$.}
\es{Los agentes exhiben distintos desempe\~nos en cada evento, que var\'ian alrededor de su verdadera habilidad, $\N(p\,|\,s,\beta^2)$.}
%
%
\en{The model assumes that the agent with the highest performance wins, $r = (p_i > p_j)$. In other words, whoever obtains a difference of performance greater than 0 wins, $r = (p_i - p_j > 0)$.}
\es{El modelo supone que gana el agente con mayor rendimiento, $r = (p_i > p_j)$, o en otras palabras, gana quien obtenga una diferencia de rendimiento mayor a 0, $r = (p_i - p_j > 0)$.}
\en{The parameter $\beta^2$, being the same for all the agents, acts as the scale of the estimations: skills separated by a $\beta$ keep always the same probability of winning.}
\es{El parámetro $\beta^2$, al ser el mismo para todos los agentes, actúa como la escala de las estimaciones: habilidades separada por un $\beta$ mantienen siempre la misma probabilidad de ganar.}

@@

\en{However, it does have a major weaknesses.}

@@

\en{The prior quantifies the uncertainty about the ability, and the posterior contains the uncertainty that remains after seeing the new data (the result of the event).}
\es{El prior cuantifica la incertidumbre sobre la habilidad, y el posterior contiene la incertidumbre que queda luego de ver el nuevo dato (el resultado del evento).}

@@

\en{The difference of performances, $d=p_i-p_j$, can also be expressed as a Gaussian distribution, centered on the difference of the prior estimate means ($\mu_i -\mu_j$), with a variance that incorporates the uncertainty of both estimates ($\sigma$) and the variance of both performances ($\beta$), $\N( d \, | \, \mu_i -\mu_j \, ,\ 2\beta^2 + \sigma_i^2 + \sigma_j^2 \,)$.}

@@

\en{From now on the role of the model will be left implicit.}

@@

\en{The likelihood quantifies the same cumulative density as the evidence, but now centered at the difference between the hypothesis we are evaluating $s_i$ and the opponent's mean estimate $\mu_j$, with a variance that includes all uncertainties except the one of $s_i$.}

@@

\en{It is important to stress that the posterior, although similar, is not a Gaussian distribution, preventing us from using equation~\ref{eq:posterior_win} iteratively.}
\es{Es importante remarcar que la posterior, aunque se parezca, no es una distribuci\'on gaussiana, lo que nos impedir\'a usar la ecuaci\'on~\ref{eq:posterior_win} iterativamente.}
%
\en{But due to the shape of the exact posterior, a Gaussian distribution could be used as a good approximation, allowing us to avoid the computational cost of the sampling methodologies.}
\es{Por la forma del posterior exacto, una gaussiana puede ser usada como una buena aproximaci\'on, permiti\'endonos evitar el costo computacional de las metodolog\'ias de sampleo.}
%
\en{The main contribution of the Glicko system~\citep{glikman_gliko_2} was the development of an efficient method to approximate the exact posterior using a Gaussian distribution.}
\es{El principal aporte del sistema Glicko~\citep{glikman_gliko_2} fue el desarroll\'o de un m\'etodo eficiente para aproximar la posterior exacta con una distribuci\'on Gasussiana.}
%
\en{However, this method does not guarantee the quality of the approximation used.}
\es{Sin embargo, este m\'etodo no garantiza que la distribuci\'on gaussiana seleccionada sea la que mejor aproxima.}
%
\en{The success of the TrueSkill solution~\citep{Herbrich2007} is based on the application of an efficient method for computing the Gaussian distribution that best approximates the exact posterior (section~\ref{sec:approximate_posterior}).}
\es{El \'exito de la soluci\'on TrueSkill~\citep{Herbrich2007} se basa en la aplicaci\'on de m\'etodo eficiente para calcular la gaussiana que mejor aproxima a la posterior exacta (secci\'on~\ref{sec:approximate_posterior}).}
%
\begin{equation} \label{eq:approx} 
 \widehat{p}(s_i| r, s_j) = \underset{\mu, \sigma}{\text{ arg min }} \ \ \text{KL}(\, p(s_i| r, s_j) \, || \,  \N(s_i|\mu, \sigma^2) \, )
\end{equation}
%
\en{This kind of methods, which minimize the Kullback-Leibler divergence between the true and the approximate distribution, allows to perform Bayesian inference in situations that otherwise would be impossible.}
\es{Esta clase de m\'etodos, que minizan la divergencia Kullback-Leibler entre la distribuci\'on verdadera y aproximada, permiten realizar inferencia bayesiana en situaciones que de otra forma no ser\'ia posibles.}
%
\en{In our case, in which we need to know how skills change over time, this technique allows us to efficiently apply the equation~\ref{eq:posterior_win} iteratively over a sequence of observations.}
\es{En nuestro caso, en el que necesitamos conocer c\'omo cambian las habilidades a trav\'es del tiempo, est\'a t\'ecnica nos permite aplicar eficientemente la ecuaci\'on~\ref{eq:posterior_win} iterativamente sobre una secuencia de observaciones.}

\en{The approach adopted by TrueSkill to treat the dynamical process, known as \emph{filtering}, uses the posterior as the prior for the next event.}
\es{El enfoque adoptado por TrueSkill para tratar el proceso din\'amico, conocido como \emph{filtering}, usa el posterior como prior del siguiente evento.}
%
\en{Then, the posterior at any given time is,}
\es{Luego, el posterior en un determinado momento es,}
%
\begin{equation}\label{eq:filter} %\tag{\text{filtering}}
 \widehat{\text{Posterior}}_t \propto \widehat{\text{Likelihood}}_t  \overbrace{\widehat{\text{Likelihood}}_{t-1} \dots \widehat{\text{Likelihood}}_{2} \underbrace{\widehat{\text{Likelihood}}_{1} \text{Prior}_1}_{\widehat{\text{Posterior}}_{1} \text{ \en{as}\es{como} } \text{Prior}_{2}} }^{\widehat{\text{Posterior}}_{t-1} \text{ \en{as}\es{como} } \text{Prior}_{t}} %= \text{Prior}_1 \prod_{i=1}^t \text{Likelihood}_i 
\end{equation}
%
\en{where {\footnotesize $\widehat{\text{Posterior}}_i$} and {\footnotesize $\widehat{\text{Likelihood}}_i$} represents the approximations induced by the equation~\ref{eq:approx} at the $i$-est event.}
\es{Donde {\footnotesize $\widehat{\text{Posterior}}_i$} y {\footnotesize $\widehat{\text{Likelihood}}_i$} representan la aproximaciones inducidas por la ecuaci\'on~\ref{eq:approx} en el $i$-\'esimo evento.}
%
\en{If we consider the likelihood as a filter of the prior, each posterior can be seen as an accumulation of all previous filters.}
\es{Si consideramos la verosimilitud como un filtro del prior, cada posterior puede ser visto como una acumulaci\'on de todos los filtros anterios.}
%
\en{In this way, information propagates from past to future estimates.}
\es{De esta forma, la informaci\'on propaga del estimaciones pasadas hacia futuras.}
%
%\en{Learning is a dynamical process.}
%\es{El aprendizaje es un proceso din\'amico.}
%
\en{Since skills are dynamic variables it is important to add some uncertainty $\gamma$ at each step.}
\es{Debido a que las habilidades son variables din\'amicas es importante agregar alguna incertidumbre $\gamma$ en cada paso.}
%
\begin{equation}\label{eq:dynamic_factor}
 \widehat{p}(s_{i_t}) = \N(s_{i_t} | \mu_{i_{t-1}}, \sigma_{i_{t-1}}^2 + \gamma^2 )
 \end{equation}
 %
\en{Because the filtering approach does not arise from any probabilistic model, it suffers from a number of problems related to the fact that the information propagates in only one direction through the system.}
\es{Debido a que el enfoque de filtrado no se surge de ningún modelo probabil\'istico, sufre de un serie de problemas, todos relacionados con el hecho de que la informaci\'on se propaga en una sola direcci\'on a trav\'es del sistema.}
%
\en{The most obvious is that the beginning of any sequence of estimates always has high uncertainty.}
%, making them useless
\es{El m\'as obvio es que el inicio de toda secuencia de estimaciones siempre tiene alta incertidumbre.}
% , lo que las hace inservibles
\en{But it also suffer from ``temporary'' and ``spatial'' decoupling.}
\es{Pero tambi\'en sufre de desacoplamientos ``temporales'' y ``espaciales''.}
%
\en{Although the relative differences between contemporary estimates within well-connected communities are correct, estimates separated in time and between poorly connected communities are often incorrect.}
% preventing comparability over time and between communities.
\es{Aunque la diferencias relativas entre estimaciones contempor\'aneas al interior de comunidades bien conectadas sean correctas, las estimaciones separadas en el tiempo y entre comunidades poco conectadas suelen ser incorrectas.}
% impidiendo as\'i la comparabilidad temporal y entre comunidades.

@@

\en{This produce a mutual dependency between likelihoods, forcing us to implement an iterative algorithm to solve inference (details at secci\'on~\ref{sec:throguthTime}).}

@@

\en{In the first pass, the priors of each event are calculated using only the past likelihoods that have already been defined.}

@@

\en{Once all the likelihoods have been calculated, we can run a new pass using the latest available likelihoods.}

@@

\en{This procedure converge with a few linear iterations over the data.}

@@

      }  
  }
  \caption{
  \en{Convergence of TrueSkill Through Time in a history with only two events and two agents.}
  \es{Convergencia de TrueSkill Through Time en una historia con s\'olo dos eventos y dos agentes.}
  %
  \en{The first game is won by the player $a$, $p_{a_1} > p_{b_1}$, and the second is won by the player $b$, $p_{a_2} < p_{b_2}$.}
  \es{La primera partida la gana el jugador $a$, $p_{a_1} > p_{b_1}$, y la segunda la gana el jugador $b$, $p_{a_2} < p_{b_2}$.}
  %
  \en{The brightness of the curves indicates the order of convergence: the darkest are the latest estimates.}
  \es{La luminosidad de las curvas indican el orden en la convergencia.}
  %
  \en{The first one (the clearest) corresponds to the TrueSkill solution.}
  \es{La primera (la m\'as clara) se correponde a la soluci\'on TrueSkill.}
  %
  \en{The last one (the darkest) corresponds to the TrueSkill Through Time solution.}
  \es{La última (la m\'as oscura) se correponde a la soluci\'on TrueSkill Through Time.}
  %
  \en{Once convergence is reached, the posterior is centered at 0 for both players, indicating that they have the same skill.}
  \es{Alcanzada la convergencia, el posterior se centra en 0 para ambos jugadores, indicando que tienen misma habilidad.}
  }
  \label{fig:smooth_example}
\end{figure}
%
\en{TrueSkill Through Time recovers the true differences between hidden skills even at the beginning of the sequence.}
\es{TrueSkill Through Time recupera las verdaderas diferencias entre habilidades ocultas incluso al principio de la secuencia.}
%  
\en{Similar methodologies to TrueSkill Through Time have been developed by \cite{coulom2008-wholeHistoryRating} and \cite{maystre2019-pairwise}.}
\es{Metodologías similares a TrueSkill Through Time han sido desarrollads por \cite{coulom2008-wholeHistoryRating} y \cite{maystre2019-pairwise}.}
%
%\en{This methodology has not been available until now in any of the major programming languages.}
%\es{Esta metodolog\'ia no hab\'ia estado hasta ahora disponible en ninguno de los principales lenguajes de programaci\'on.}
%
\en{With this paper we make available the first TrueSkill Through Time packages for \proglang{Julia}, \proglang{Python} and \proglang{R}, together with its scientific documentation.}
\es{Con este art\'iculo ponemos a disposici\'on los primeros paquetes de TrueSkill Through Time para \proglang{Julia}, \proglang{Python} y \proglang{R}, junto con su documentaci\'on cient\'ifica.}

>>>>

\en{Knowing how individual skills change over time is essential in Sports and Education.}
\es{Conocer c\'omo cambian las habilidades individuales a lo largo del tiempo es esencial para en contextos deportivos y educativos.}
%
\en{Since the skill is a hidden variable, the best we can do is to estimate it based on its direct observable consequences: the outcome of problem solving and competitions.}
\es{Dado que son variables ocultas, lo mejor que podemos hacer es estimarlas a partir de sus consecuencias observables directas: el producto de resoluci\'on de problemas y competencias.}
%
\en{Considering only the frequency of positive results as an indicator of the individuals' ability could lead to wrong approximations, mainly because the outcome also depends on the difficulty of the challenge.}
\es{Considerar s\'olo la frecuencia de resultados positivos como indicador de la habilidad de los individuos puede conducir a aproximaciones erroneas, fundamentalmente porque su valor depende tambi\'en de la dificultad de los desaf\'ios.}
%
\en{For this reason, all widely used skill estimators are based on pairwise comparisons.}
\es{Por esta raz\'on, todos los estimadores de habilidad ampliamente usados se basan en comparaciones por pares.}
%
\en{Since the first generative models, proposed almost a century ago by~\cite{Thurstone1927} and~\cite{Zermelo1929}, it is assumed that the probability of the observed result $r$ depends on the performance $p$ of the agent $i$ and their opponent $j$, expressed as $P(\, r \,|\, p_i, \, p_j \,)$.}
\es{Desde los primeros modelos generativos, propuestos hace casi un siglo por~\cite{Thurstone1927} y~\cite{Zermelo1929}, se supone que la probabilidad de un resultado observado, $r$, depende del rendimiento, $p$, del agente $i$ y de su oponente $j$, expresada como $P(\, r \,|\, p_i, \, p_j \,)$.}
%
\en{The field continued to progress with the work of~\cite{Bradley1952} and~\cite{Mosteller1951a,Mosteller1951b,Mosteller1951c}, leading to the breakthrough that took place with the methodology developed by~\cite{Elo2008} for the US Chess Federation (USCF), which is still used by the International Chess Federation (FIDE).}
\es{El campo sigui\'o progresando con los trabajos de \cite{Bradley1952} y~\cite{Mosteller1951a,Mosteller1951b,Mosteller1951c}, que condujo al gran avance que tuvo lugar con la metodología desarrollada por~\cite{Elo2008} para la Federaci\'on de Ajedrez de los Estados Unidos (USCF), adoptada hasta el d\'ia de hoy por la Federaci\'on Internacional de Ajedrez (FIDE).}

@@

\en{All currently used skill estimators share some variant of the probabilistic model proposed by Elo~\citep{Glickman1999,Herbrich2007,VanDerLinden2016,Fox2010}.}
\es{Todos los estimadores de habilidad ampliamente utilizados actualmente comparten alguna variante del modelo probabil\'istico propuesto por Elo~\cite{Glickman1999, Herbrich2007, VanDerLinden2016, Fox2010}.}
%
\begin{figure}[h!]
\centering \small
    \tikz{         
    \node[det, fill=black!10] (r) {$r$} ; 
    \node[const, left=of r, xshift=-1.35cm] (r_name) {\small \en{Result}\es{Resultado}:}; 
    \node[const, right=of r] (dr) {\normalsize $ r = (d > 0)$};

@@

    \edge {d} {r};
    \edge {p1,p2} {d};
    \edge {s1} {p1};
    \edge {s2} {p2};
    %\node[invisible, right=of p2, xshift=4.35cm] (s-dist) {};
}
     \caption{
     \en{Generative model in which skills cause the observable results mediated by the difference of hidden performances, $d =p_i - p_j$, both random variables around their unknown true skill, $p \sim \N(s,\beta^2)$.}
    \es{Modelo generativo en el que las habilidades causan los resultados observables a trav\'es de la diferencia de rendimientos ocultos, $d=p_i-p_j$, ambas variables aleatorias centradas en la verdadera habilidad, $p \sim \N(s,\beta^2)$}
    %
    \en{The one with the highest performance wins, $r = (d > 0)$.}
    \es{Gana quien haya obtenido mayor rendimiento, $r = (d > 0)$.}
    %
    \en{Observable variables are painted gray, hidden in white, and constants are shown as black dots.}
    \es{Las variables observables se pintan de gris, la ocultas en blanco, y las constantes se muestran como puntos negros.}
    }
    \label{fig:generative_model}
\end{figure}
%
\en{Figure~\ref{fig:generative_model} provides a causal model in which skills generates the observable result.}
\es{La figura~\ref{fig:generative_model} ofrece un modelo causal en la que las habilidades generan el resultado observable.}
%
\en{The agents exhibit different performances at each event, varying around their true skill, $\N(p_i\,|\,s_i,\beta^2)$.}
\es{Los agentes exhiben distintos desempe\~nos en cada evento, que var\'ian alrededor de su verdadera habilidad, $\N(p\,|\,s,\beta^2)$.}
%
%
\en{The model assumes that the agent with the highest performance wins, $r = (p_i > p_j)$. In other words, whoever obtains a difference of performance greater than 0 wins, $r = (p_i - p_j > 0)$.}
\es{El modelo supone que gana el agente con mayor rendimiento, $r = (p_i > p_j)$, o en otras palabras, gana quien obtenga una diferencia de rendimiento mayor a 0, $r = (p_i - p_j > 0)$.}
\en{The parameter $\beta^2$ is the same for all the agents and acts as the scale of the estimations: skills separated by a $\beta$ keep always the same probability of winning.}
\es{El parámetro $\beta^2$, al ser el mismo para todos los agentes, actúa como la escala de las estimaciones: habilidades separada por un $\beta$ mantienen siempre la misma probabilidad de ganar.}

@@

\en{However, it does have a major weakness.}

@@

\en{The prior quantifies the uncertainty about the ability, and the posterior contains the uncertainty that remains after seeing the new data (i.e.~the result of the event).}
\es{El prior cuantifica la incertidumbre sobre la habilidad, y el posterior contiene la incertidumbre que queda luego de ver el nuevo dato (i.e.~el resultado del evento).}

@@

\en{The difference of performances, $d=p_i-p_j$, can also be expressed as a Gaussian distribution, centered on the difference of the prior estimate means ($\mu_i -\mu_j$), with a variance that incorporates the uncertainty of both estimates ($\sigma_i$ and $\sigma_j$) and the variance of both performances ($\beta$), $\N( d \, | \, \mu_i -\mu_j \, ,\ 2\beta^2 + \sigma_i^2 + \sigma_j^2 \,)$.}

@@

\en{During the rest of this work, the role of the model will be left implicit.}

@@

\en{The likelihood quantifies the same cumulative density as the evidence but centered at the difference between the hypothesis we are evaluating $s_i$ and the opponent's mean estimate $\mu_j$, with a variance that includes all uncertainties except the one of $s_i$.}

@@

\en{It is important to stress that the posterior, although similar, is not a Gaussian distribution, preventing us from using equation~\ref{eq:posterior_win} iteratively.}
\es{Es importante remarcar que la posterior, aunque se parezca, no es una distribuci\'on gaussiana, lo que nos impedir\'a usar la ecuaci\'on~\ref{eq:posterior_win} iterativamente.}
%
\en{But due to the shape of the exact posterior, a Gaussian distribution could be used as a good approximation, allowing us to avoid the computational cost of the sampling methodologies.}
\es{Por la forma del posterior exacto, una gaussiana puede ser usada como una buena aproximaci\'on, permiti\'endonos evitar el costo computacional de las metodolog\'ias de sampleo.}
%
\en{The main contribution of the Glicko system~\citep{glikman_gliko_2} was the development of an efficient method to approximate the exact posterior using a Gaussian distribution.}
\es{El principal aporte del sistema Glicko~\citep{glikman_gliko_2} fue el desarroll\'o de un m\'etodo eficiente para aproximar la posterior exacta con una distribuci\'on Gasussiana.}
%
\en{However, this method does not guarantee the quality of the approximation used.}
\es{Sin embargo, este m\'etodo no garantiza que la distribuci\'on gaussiana seleccionada sea la que mejor aproxima.}
%
\en{The success of the TrueSkill solution~\citep{Herbrich2007} is based on the application of an efficient method for computing the Gaussian distribution that guarantees the approximation of the posterior (section~\ref{sec:approximate_posterior}).}
\es{El \'exito de la soluci\'on TrueSkill~\citep{Herbrich2007} se basa en la aplicaci\'on de m\'etodo eficiente para calcular la gaussiana que mejor aproxima a la posterior exacta (secci\'on~\ref{sec:approximate_posterior}).}
%
\begin{equation} \label{eq:approx} 
 \widehat{p}(s_i| r, s_j) = \underset{\mu, \sigma}{\text{ arg min }} \ \ \text{KL}(\, p(s_i| r, s_j) \, || \,  \N(s_i|\mu, \sigma^2) \, )
\end{equation}
%
\en{which minimize the Kullback-Leibler divergence between the true and the approximate distribution. This  kind of methods allows performing Bayesian inference in situations that otherwise would be impossible.}
\es{Esta clase de m\'etodos, que minizan la divergencia Kullback-Leibler entre la distribuci\'on verdadera y aproximada, permiten realizar inferencia bayesiana en situaciones que de otra forma no ser\'ia posibles.}
%
\en{In our case, in which we need to know how skills change over time, this technique allows us to efficiently apply equation~\ref{eq:posterior_win} iteratively over a sequence of observations.}
\es{En nuestro caso, en el que necesitamos conocer c\'omo cambian las habilidades a trav\'es del tiempo, est\'a t\'ecnica nos permite aplicar eficientemente la ecuaci\'on~\ref{eq:posterior_win} iterativamente sobre una secuencia de observaciones.}

\en{The approach adopted by TrueSkill to treat the dynamical process, known as \emph{filtering}, uses the posterior as the prior for the next event.}
\es{El enfoque adoptado por TrueSkill para tratar el proceso din\'amico, conocido como \emph{filtering}, usa el posterior como prior del siguiente evento.}
%
\en{Then, the posterior at any given time is,}
\es{Luego, el posterior en un determinado momento es,}
%
\begin{equation}\label{eq:filter} %\tag{\text{filtering}}
 \widehat{\text{Posterior}}_t \propto \widehat{\text{Likelihood}}_t  \overbrace{\widehat{\text{Likelihood}}_{t-1} \dots \widehat{\text{Likelihood}}_{2} \underbrace{\widehat{\text{Likelihood}}_{1} \text{Prior}_1}_{\widehat{\text{Posterior}}_{1} \text{ \en{as}\es{como} } \text{Prior}_{2}} }^{\widehat{\text{Posterior}}_{t-1} \text{ \en{as}\es{como} } \text{Prior}_{t}} %= \text{Prior}_1 \prod_{i=1}^t \text{Likelihood}_i 
\end{equation}
%
\en{where {\footnotesize $\widehat{\text{Posterior}}_i$} and {\footnotesize $\widehat{\text{Likelihood}}_i$} represents the approximations induced by the equation~\ref{eq:approx} at the $i$-th event.}
\es{Donde {\footnotesize $\widehat{\text{Posterior}}_i$} y {\footnotesize $\widehat{\text{Likelihood}}_i$} representan la aproximaciones inducidas por la ecuaci\'on~\ref{eq:approx} en el $i$-\'esimo evento.}
%
\en{If we consider the likelihood as a filter of the prior, each posterior can be seen as an accumulation of all previous filters.}
\es{Si consideramos la verosimilitud como un filtro del prior, cada posterior puede ser visto como una acumulaci\'on de todos los filtros anterios.}
%
\en{In this way, information propagates from past to future estimates.}
\es{De esta forma, la informaci\'on propaga del estimaciones pasadas hacia futuras.}
%
%\en{Learning is a dynamical process.}
%\es{El aprendizaje es un proceso din\'amico.}
%
\en{Since skills are dynamic variables it is important to add some uncertainty $\gamma$ at each step.}
\es{Debido a que las habilidades son variables din\'amicas es importante agregar alguna incertidumbre $\gamma$ en cada paso.}
%
\begin{equation}\label{eq:dynamic_factor}
 \widehat{p}(s_{i_t}) = \N(s_{i_t} | \mu_{i_{t-1}}, \sigma_{i_{t-1}}^2 + \gamma^2 )
 \end{equation}
 %
\en{Because the filtering approach does not arise from any probabilistic model, it suffers from some problems related to the fact that the information propagates in only one direction through the system.}
\es{Debido a que el enfoque de filtrado no se surge de ningún modelo probabil\'istico, sufre de un serie de problemas, todos relacionados con el hecho de que la informaci\'on se propaga en una sola direcci\'on a trav\'es del sistema.}
%
\en{The most obvious is that the beginning of any sequence of estimates always has high uncertainty.}
%, making them useless
\es{El m\'as obvio es que el inicio de toda secuencia de estimaciones siempre tiene alta incertidumbre.}
% , lo que las hace inservibles
\en{But it also suffers from ``temporary'' and ``spatial'' decoupling.}
\es{Pero tambi\'en sufre de desacoplamientos ``temporales'' y ``espaciales''.}
%
\en{Although the relative differences between current estimates within well-connected communities are correct, estimates separated in time and between poorly connected communities are often incorrect.}
% preventing comparability over time and between communities.
\es{Aunque la diferencias relativas entre estimaciones contempor\'aneas al interior de comunidades bien conectadas sean correctas, las estimaciones separadas en el tiempo y entre comunidades poco conectadas suelen ser incorrectas.}
% impidiendo as\'i la comparabilidad temporal y entre comunidades.

@@

\en{This procedure establishes a mutual dependency between likelihoods, forcing us to implement an iterative algorithm to solve the inference (details in section~\ref{sec:throguthTime}).}

@@

\en{In the first pass, we use the past likelihoods to compute the priors of each event.}

@@

\en{Once all the likelihoods have been obtained, we can perform a new iteration using the recently computed likelihoods.}

@@

\en{This procedure converges with a few linear iterations over the data.}

@@

      }  
  }
  \caption{
  \en{The convergence of TrueSkill Through Time in a history consisting of two events and two agents.}
  \es{Convergencia de TrueSkill Through Time en una historia con s\'olo dos eventos y dos agentes.}
  %
  \en{The first game is won by player $a$, $p_{a_1} > p_{b_1}$, and the second one is won by player $b$, $p_{a_2} < p_{b_2}$.}
  \es{La primera partida la gana el jugador $a$, $p_{a_1} > p_{b_1}$, y la segunda la gana el jugador $b$, $p_{a_2} < p_{b_2}$.}
  %
  \en{The brightness of the curves indicates the order of convergence: the darkest are the latest estimates.}
  \es{La luminosidad de las curvas indican el orden en la convergencia.}
  %
  \en{The first one (the clearest) corresponds to the TrueSkill solution.}
  \es{La primera (la m\'as clara) se correponde a la soluci\'on TrueSkill.}
  %
  \en{The last one (the darkest) corresponds to the TrueSkill Through Time solution.}
  \es{La última (la m\'as oscura) se correponde a la soluci\'on TrueSkill Through Time.}
  %
  \en{Once convergence is reached, the posterior is centered at 0 for both players, indicating that they have the same skill.}
  \es{Alcanzada la convergencia, el posterior se centra en 0 para ambos jugadores, indicando que tienen misma habilidad.}
  }
  \label{fig:smooth_example}
\end{figure}
%
\en{TrueSkill Through Time recovers the actual differences between hidden skills even at the beginning of the sequence.}
\es{TrueSkill Through Time recupera las verdaderas diferencias entre habilidades ocultas incluso al principio de la secuencia.}
%  
\en{Similar methodologies to TrueSkill Through Time have been developed by \cite{coulom2008-wholeHistoryRating} and \cite{maystre2019-pairwise}.}
\es{Metodologías similares a TrueSkill Through Time han sido desarrollads por \cite{coulom2008-wholeHistoryRating} y \cite{maystre2019-pairwise}.}
%
%\en{This methodology has not been available until now in any of the major programming languages.}
%\es{Esta metodolog\'ia no hab\'ia estado hasta ahora disponible en ninguno de los principales lenguajes de programaci\'on.}
%
\en{Jointly with this work, we make available the first TrueSkill Through Time packages for \proglang{Julia}, \proglang{Python}, and \proglang{R}, together with its scientific documentation.}
\es{Con este art\'iculo ponemos a disposici\'on los primeros paquetes de TrueSkill Through Time para \proglang{Julia}, \proglang{Python} y \proglang{R}, junto con su documentaci\'on cient\'ifica.}
